An Information Lens on Demographic Diversity

### 2016-10-06 5_figs.Rmd

```{r, fig.height=3}
checkerboard_illustration()
ggsave('figs/checkerboard.png', width = 4, height = 1.5)
```

@@

first <- method_illustration('Detroit', vmax = 6)
second <- method_illustration('Philadelphia', vmax = 6)

gA <- ggplotGrob(first)
gB <- ggplotGrob(second)
g <- gridExtra::cbind.gtable(gA, gB, size = "max")

grid.draw(g)

ggsave('figs/method_illustration.png', g, width = 8, height = 5)

@@

ggsave('figs/decomposability.png', width = 2.5, height = 1.5)

@@

cache %>%
	ggplot(aes(x = I_XY, y = weighted_J)) + 
	geom_point(color = 'grey') + 
	theme_minimal() + 
	ylab(expression(J(X,Y))) + 
	xlab(expression(I(X,Y))) + 
	ggtitle('Mutual information and mean local information in major US cities') +
	geom_point(data = highlight, color = 'firebrick') + 
	geom_text_repel(data = highlight, aes(label = city))

ggsave('figs/mutual_fisher.png', width = 8, height = 4)

@@

## Relationship to population density
```{r}
lm_eqn <- function(df){
	m <- lm(weighted_J ~ log(density), data = df) 
	eq <- substitute(italic(y) == a + b %.% ~'log' ~italic(rho)*","~~italic(r)^2~"="~r2, 
					 list(a = format(coef(m)[1], digits = 2), 
					 	 b = format(coef(m)[2], digits = 2), 
					 	 r2 = format(summary(m)$adj.r.squared, digits = 3)))
	as.character(as.expression(eq))                 
}

cache %>% 
	# filter(!grepl('Los Angeles', city)) %>%
	ggplot(aes(x = density, y = weighted_J)) +
	geom_smooth(method = 'lm', se = FALSE, color = 'black', size = .4) +
	annotate('text', x = 3000, y = .2, label = lm_eqn(cache), parse = T) + 
	geom_point(color = 'grey') + 
	geom_point(color = 'firebrick', data = highlight) + 
	geom_text_repel(aes(label = city), size = 3, data = highlight) +
	scale_x_continuous(trans = 'log10') +
	# scale_y_continuous(trans = 'log10') +
	theme_minimal() +
	xlab(expression(rho~(population/km^{2}))) +
	ylab(expression(J(X,Y))) +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Dependence of spatial complexity on population density')

ggsave('figs/density_fisher.png', width = 8, height = 4)
```

@@

tab <- cache %>% 
	select(area, population, H_Y, I_XY, J = weighted_J) %>% 
	mutate(Density = population / area) %>% 
	round(2) %>%
	cbind(cache$city, .) %>%
	rename(City = `cache$city`, 
		   `Area` = area,
		   Population = population,
		   HY = H_Y,
		   IXY = I_XY) %>%
	arrange(desc(Density)) %>% 
	select(City, `Area`, Population, Density, HY, IXY, J)

@@

join_cols <- c('area', 'population', 'density', 'H_X', 'H_Y', 'local.H_Y', 'weighted_J', 'city')
loss_curves <- loss_curves %>% left_join(info_summary[,join_cols])

@@

ggsave('figs/example_cluster_map.png',  width = 8, height = 4)

@@

	I <- mutual_info(tracts@data[,c(columns, 'cluster')],group_col = 'cluster')

@@

ggsave('figs/example_clusters.png')

@@

ggsave('figs/example_clusters_detailed.png', width = 4, height = 2)

@@

```{r}
end_points <- loss_curves %>%
	group_by(city) %>%
	filter(nclust == max(nclust)) %>% 
	filter(weighted_J > 1.25 | weighted_J < .75)

loss_curves %>% 
	ggplot() +
	aes(x = nclust, y = I_XY, group = city) +
	geom_path(aes(color = weighted_J)) +
	# geom_point(aes(color = weighted_J), size = .3) +
	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
	# scale_y_continuous(trans = 'log10') +
	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
	theme_minimal() +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')

ggsave('figs/loss_curves.png', width = 8, height = 4)

@@

loss_curves %>% 
	group_by(city) %>% 
	mutate(I_max = max(I_XY)) %>% 
	ggplot() +
	aes(x = nclust, y = I_XY, group = city) +
	geom_path(aes(color = I_max)) +
	# geom_point(aes(color = weighted_J), size = .3) +
	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
	# scale_y_continuous(trans = 'log10') +
	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
	theme_minimal() +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')

@@

loss_curves %>%
	group_by(city) %>%
	mutate(I_XY = I_XY / max(I_XY)) %>%
	mutate(nclust = nclust / max(nclust)) %>%
	# filter(nclust < 100) %>%
	ggplot() +
	aes(x = nclust, y = I_XY, color = weighted_J, group = city) +
	geom_path() +
	scale_x_continuous(trans = 'log10') +
	scale_y_continuous(trans = 'log10') +
	theme_minimal() +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	annotation_logticks(size = .1, sides = 'b') + 
	ggtitle('Normalized loss curves for spatially-constrained information-theoretic clustering')

ggsave('figs/normed_loss_curves.png', width = 8, height = 4)
```

@@

compute_AUC <- function(loss_curves, adjust_thresh = 0){
	loss_curves %>% 
		group_by(city) %>%
		mutate(exclude = (nclust != 1) & (cummax(I_XY - lag(I_XY,default = 0)) < adjust_thresh) & (nclust < 10)) %>% 
		filter(!exclude) %>%
		mutate(nclust = row_number()) %>% 
		mutate(dx = log(nclust / (nclust - 1))) %>%
		filter(nclust > 1) %>%
		group_by(city) %>%
		arrange(nclust) %>%
		summarise(corner = max(I_XY) * log(max(nclust)),
				  area = trapz(nclust, I_XY / nclust)) %>%
		mutate(AUC = area / corner) %>%
		arrange(desc(AUC)) 
}

summary <- compute_AUC(loss_curves, adjust_thresh = 0)

summary <- summary %>% left_join(cache, by = c('city' = 'city'))

# %>%
# 	left_join(info_summary[,c('city', 'weighted_J', 'I_XY')]) %>%
# 	filter(I_XY > 0)

@@

filled <- sub_curves[,c('nclust','I_XY', 'city')]

corner <- filled %>% group_by(city) %>%
	summarise(nclust = max(nclust),
			  I_XY = 0)

filled <- rbind(corner, filled) %>%
	arrange(city, nclust, desc(I_XY))

sub_curves %>% 
	ggplot() +
	geom_path(aes(x = nclust, y = I_XY, group = city, color = city)) +
	scale_x_continuous(trans = 'log10', limit = c(1, NA)) +
	scale_y_continuous(limit = c(0,max(sub_curves$I_XY))) +
	theme_minimal() +
	annotation_logticks(size = .1, sides = 'b', scaled = T) + 
	geom_rect(data = rects, aes(xmin = xmin,
								ymin = ymin,
								xmax = xmax,
								ymax = ymax,
								group = city),
			  color = 'grey',
			  alpha = 0) +
	# geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) + 
	geom_polygon(aes(nclust, I_XY, fill = city), data = filled, alpha = .2) + 
	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(weighted_J,2))), data = sub_summary) +
	facet_wrap(~city) + 
	scale_fill_discrete(guide = F) + 
	scale_color_discrete(guide = F) + 
	xlab('Number of clusters') + 
	ylab('I(C,Y)') + 
	ggtitle('AUC as Clustering Evaluation')

ggsave('figs/AUC_illustration.png', width = 8, height = 4)
```

@@

summary[,c('I_XY', 'weighted_J', 'AUC', 'adj_AUC')] %>% 
	round(2) %>% 
	cbind(summary$city) %>% 
	select(city = `summary$city`, I_XY, weighted_J, AUC, adj_AUC) %>%
	datatable(options = list(pageLength = 10), rownames = FALSE,
		  caption = 'Table 1: This is a simple caption for the table.')
```

@@

summary %>%
	ggplot() +
	aes(x = I_XY, y = AUC) +
	geom_point(aes(color = weighted_J)) +
	geom_smooth(color = 'grey', se = F) +
	geom_text_repel(aes(label = city), data = highlight) + 
	theme_minimal() + 
	ggtitle('Clustering performance and mutual information') +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	xlab('I(X,Y)') + 
	ylab('AUC')

ggsave('figs/info_performance.png', width = 8, height = 4)

@@

summary %>%
	ggplot() +
	aes(x = I_XY, y = adj_AUC) +
	geom_point(aes(color = weighted_J)) +
	geom_smooth(color = 'grey', se = F) +
	geom_text_repel(aes(label = city), data = highlight) + 
	theme_minimal() + 
	ggtitle('Clustering performance and mutual information') +
	xlab('I(X,Y)') + 
	ylab('AUC (geographically adjusted)') + 
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,.2), name = "J(X,Y)", guide = guide_colorbar(direction = 'horizontal')) +
	theme(legend.justification=c(1.5,0),
			  legend.position=c(1,0))

ggsave('figs/info_performance_adjed.png', width = 8, height = 4)

@@

```{r}
summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
		   `I(X,Y)` = I_XY / max(I_XY),
		   `J(X,Y)` = weighted_J / max(weighted_J),
		   `H(Y)` = H_Y / max(H_Y)) %>%
	filter(I_XY > 0) %>%
	lm(AUC ~ `I(X,Y)` + `J(X,Y)`, data = .) %>%
	stargazer(flip = T, 
			  # table.layout = '#tn', 
			  style = 'all2',
			  ci = T,covariate.labels = c('$I(X,Y)$','$J(X,Y)$', 'Intercept'),
			  table.layout = '-=-!=!dtas',
			  float = F) %>% 
	write(file = 'figs/adjusted_regression.txt')
```

>>>>

```{r, fig.height=3}
checkerboard_illustration()
ggsave('figs/checkerboard.pdf', width = 4, height = 1.5)
```

@@

# first <- method_illustration('Detroit', vmax = 6)
# second <- method_illustration('Philadelphia', vmax = 6)
# 
# gA <- ggplotGrob(first)
# gB <- ggplotGrob(second)
# g <- gridExtra::cbind.gtable(gA, gB, size = "max")
# 
# grid.draw(g)
# 
# ggsave('figs/method_illustration.pdf', g, width = 8, height = 5)

@@

ggsave('figs/decomposability.pdf', width = 2.5, height = 1.5)

@@

cache %>%
	ggplot(aes(x = I_XY, y = J_XY)) + 
	geom_point(color = 'grey') + 
	theme_light() + 
	ylab(expression(J(X,Y))) + 
	xlab(expression(I(X,Y))) + 
	ggtitle('Mutual information and mean local information in major US cities') +
	geom_point(data = highlight, color = 'firebrick') + 
	geom_text_repel(data = highlight, aes(label = city))

ggsave('figs/mutual_fisher.pdf', width = 10, height = 6)

@@

## Relationship to population density
```{r}
lm_eqn <- function(df){
	m <- lm(J_XY ~ log(density) + I_XY, data = df) 
	eq <- substitute(italic(y) == a + b %.% 'I(X,Y)' + c %.% ~'log' ~italic(rho)*","~~italic(r)^2~"="~r2, 
					 list(a = format(coef(m)[1], digits = 2), 
					 	 b = format(coef(m)[2], digits = 2), 
					 	 c = format(coef(m)[3], digits = 2), 
					 	 r2 = format(summary(m)$adj.r.squared, digits = 3)))
	as.character(as.expression(eq))                 
}

cache %>% 
	rename(`I(X,Y)` = I_XY) %>% 
	# filter(!grepl('Los Angeles', city)) %>%
	ggplot(aes(x = density, y = J_XY)) +
	geom_smooth(method = 'lm', se = FALSE, color = 'black', size = .4) +
	annotate('text', x = 3000, y = .2, label = lm_eqn(cache), parse = T) + 
	geom_point(color = 'grey') + 
	geom_point(color = 'firebrick', data = highlight) + 
	geom_text_repel(aes(label = city), size = 3, data = highlight) +
	scale_x_continuous(trans = 'log10') +
	# scale_y_continuous(trans = 'log10') +
	theme_minimal() +
	xlab(expression(rho~(population/km^{2}))) +
	ylab(expression(J(X,Y))) +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Dependence of spatial complexity on population density')

ggsave('figs/density_fisher.pdf', width = 10, height = 6)
```

@@

tab <- cache %>% 
	select(area, population, H_Y, I_XY, J = J_XY) %>% 
	mutate(Density = population / area) %>% 
	round(2) %>%
	cbind(cache$city, .) %>%
	rename(City = `cache$city`, 
		   `Area` = area,
		   Population = population,
		   HY = H_Y,
		   IXY = I_XY) %>%
	arrange(desc(Density)) %>% 
	select(City, `Area`, Population, Density, HY, IXY, J)

@@

join_cols <- c('area', 'population', 'density',  'H_Y', 'J_XY', 'city')
loss_curves <- loss_curves %>% left_join(info_summary[,join_cols])

@@

ggsave('figs/example_cluster_map.pdf',  width = 10, height = 6)

@@

	I <- mutual_info(tracts@data[,columns])

@@

ggsave('figs/example_clusters.pdf', width = 10, height = 6)

@@

ggsave('figs/example_clusters_detailed.pdf', width = 10, height = 6)

@@

```{r}
end_points <- loss_curves %>%
	group_by(city) %>%
	filter(nclust == max(nclust)) %>% 
	filter(J_XY > 1.25 | J_XY < .75)

loss_curves %>% 
	ggplot() +
	aes(x = nclust, y = I_XY, group = city) +
	geom_path(aes(color = J_XY)) +
	# geom_point(aes(color = J_XY), size = .3) +
	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
	# scale_y_continuous(trans = 'log10') +
	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
	theme_minimal() +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')

ggsave('figs/loss_curves.pdf', width = 10, height = 6)

@@

loss_curves %>% 
	group_by(city) %>% 
	mutate(I_max = max(I_XY)) %>% 
	ggplot() +
	aes(x = nclust, y = I_XY, group = city) +
	geom_path(aes(color = I_max)) +
	# geom_point(aes(color = J_XY), size = .3) +
	scale_x_continuous(trans = 'log10', limits = c(1, max(loss_curves$nclust) + 1000)) +
	# scale_y_continuous(trans = 'log10') +
	geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) +
	theme_minimal() +
	annotation_logticks(size = .1, sides = 'b') +
	ggtitle('Loss curves for spatially-constrained information-theoretic clustering')

@@

loss_curves %>%
	group_by(city) %>%
	mutate(I_XY = I_XY / max(I_XY)) %>%
	mutate(nclust = nclust / max(nclust)) %>%
	# filter(nclust < 100) %>%
	ggplot() +
	aes(x = nclust, y = I_XY, color = J_XY, group = city) +
	geom_path() +
	scale_x_continuous(trans = 'log10') +
	scale_y_continuous(trans = 'log10') +
	theme_minimal() +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	annotation_logticks(size = .1, sides = 'b') + 
	ggtitle('Normalized loss curves for spatially-constrained information-theoretic clustering')

ggsave('figs/normed_loss_curves.pdf', width = 10, height = 6)
```

@@

compute_AUC <- function(loss_curves, adjust_thresh = 0){
	loss_curves %>% 
		group_by(city) %>%
		mutate(exclude = (nclust != 1) & (cummax(I_XY - lag(I_XY,default = 0)) < adjust_thresh) & (nclust < 10)) %>% 
		filter(!exclude) %>%
		mutate(nclust = row_number()) %>% 
		mutate(dx = log(nclust / (nclust - 1))) %>%
		filter(nclust > 1) %>%
		group_by(city) %>%
		dplyr::arrange(nclust) %>%
		summarise(corner = max(I_XY) * log(max(nclust)),
				  area = trapz(nclust, I_XY / nclust)) %>%
		mutate(AUC = area / corner) %>%
		dplyr::arrange(desc(AUC)) 
}

summary <- compute_AUC(loss_curves, adjust_thresh = 0) 
summary <- summary %>% left_join(cache, by = c('city' = 'city'))

# %>%
# 	left_join(info_summary[,c('city', 'J_XY', 'I_XY')]) %>%
# 	filter(I_XY > 0)

@@

filled <- sub_curves[,c('nclust','I_XY', 'city')] %>% 
  select(city, I_XY, nclust) %>% 
  ungroup()

corner <- filled %>% group_by(city) %>%
	summarise(I_XY = 0,
	          nclust = max(nclust))

filled <- rbind(corner, filled) %>%  # problem is right here, the columns don't add up
	dplyr::arrange(city, nclust, desc(I_XY))

sub_curves %>% 
	ggplot() +
	geom_path(aes(x = nclust, y = I_XY, group = city, color = city)) +
	scale_x_continuous(trans = 'log10', limit = c(1, NA)) +
	scale_y_continuous(limit = c(0,max(sub_curves$I_XY))) +
	theme_minimal() +
	annotation_logticks(size = .1, sides = 'b', scaled = T) + 
	geom_rect(data = rects, aes(xmin = xmin,
								ymin = ymin,
								xmax = xmax,
								ymax = ymax,
								group = city),
			  color = 'grey',
			  alpha = 0) +
	# geom_text(aes(x = nclust, y = I_XY, label = city, hjust = -.2), data = end_points) + 
	geom_polygon(aes(nclust, I_XY, fill = city), data = filled, alpha = .2) + 
	geom_text(aes(x = 50, y = .2, label = paste('AUC = ', round(AUC,2))), data = sub_summary) + 
	geom_text(aes(x = 50, y = .1, label = paste('J(X,Y) = ', round(J_XY,2))), data = sub_summary) +
	facet_wrap(~city) + 
	scale_fill_discrete(guide = F) + 
	scale_color_discrete(guide = F) + 
	xlab('Number of clusters') + 
	ylab('I(C,Y)') + 
	ggtitle('AUC as Clustering Evaluation')

ggsave('figs/AUC_illustration.pdf', width = 10, height = 6)
```

@@

summary[,c('I_XY', 'J_XY', 'AUC', 'adj_AUC')] %>% 
	round(2) %>% 
	cbind(summary$city) %>% 
	select(city = `summary$city`, I_XY, J_XY, AUC, adj_AUC) %>%
	datatable(options = list(pageLength = 10), rownames = FALSE,
		  caption = 'Table 1: This is a simple caption for the table.')
```

@@

summary %>%
	ggplot() +
	aes(x = I_XY, y = AUC) +
	geom_point(aes(color = J_XY)) +
	geom_smooth(color = 'grey', se = F) +
	geom_text_repel(aes(label = city), data = highlight) + 
	theme_minimal() + 
	ggtitle('Clustering performance and mutual information') +
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,2)) +
	xlab('I(X,Y)') + 
	ylab('AUC')

ggsave('figs/info_performance.pdf', width = 10, height = 6)

@@

summary %>%
	ggplot() +
	aes(x = I_XY, y = adj_AUC) +
	geom_point(aes(color = J_XY)) +
	geom_smooth(color = 'grey', se = F) +
	geom_text_repel(aes(label = city), data = highlight) + 
	theme_minimal() + 
	ggtitle('Clustering performance and mutual information') +
	xlab('I(X,Y)') + 
	ylab('AUC (geographically adjusted)') + 
	scale_color_continuous(low = 'white', high = 'firebrick',limits = c(0,.2), name = "J(X,Y)", guide = guide_colorbar(direction = 'horizontal')) +
	theme(legend.justification=c(1.5,0),
			  legend.position=c(1,0))

ggsave('figs/info_performance_adjed.pdf', width = 10, height = 6)

@@

```{r}
summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
		   `I(X,Y)` = I_XY / max(I_XY),
		   `J(X,Y)` = J_XY / max(J_XY),
		   `H(Y)` = H_Y / max(H_Y)) %>%
	filter(I_XY > 0) %>%
	lm(AUC ~ `H(Y)` + `I(X,Y)` + `J(X,Y)`, data = .) %>% 
	stargazer(flip = T, 
			  # table.layout = '#tn', 
			  style = 'all2',
			  ci = T,covariate.labels = c('$H(Y)$', '$I(X,Y)$','$J(X,Y)$', 'Intercept'),
			  # table.layout = '-=-!=!dtas',
			  float = F) %>% 
	write(file = 'figs/adjusted_regression.txt')
```

summary %>% mutate(AUC = adj_AUC / max(adj_AUC),
		   `I(X,Y)` = I_XY / max(I_XY),
		   `J(X,Y)` = J_XY / max(J_XY),
		   `H(Y)` = H_Y / max(H_Y),
		   density = density / max(density)) %>%
	lm(`J(X,Y)` ~ log(density) + `I(X,Y)`, data = .) %>% 
	summary()

```{r}
cities <- c('Detroit', 'Philadelphia')

tab <- data.frame(city = character(), percent_black = numeric(), H_Y = numeric(), I_XY = numeric(), J_XY = numeric())

for(city in cities){
  tracts <- readOGR(dsn = paste0('data/cities/',city), 
  	                         layer = 'geo', 
  	                         verbose = FALSE)

  grid_tract <- read_csv(file = paste0('throughput/grid_tracts/', city, '.csv'))
  grid <- readOGR(dsn = paste0('throughput/grids/',city), 
  	                         layer = 'grid', 
  	                         verbose = FALSE)

  df <- tracts@data %>%
    mutate(percent_black = Black / total) %>% 
    tbl_df

  tracts %>% 
    fortify(region = 'GEOID') %>% 
    tbl_df() %>% 
    left_join(df, by = c('id' = 'GEOID')) %>% 
    ggplot(aes(x = long, y = lat, fill = percent_black, group = group)) + 
    geom_polygon() + 
    viridis::scale_fill_viridis(name = 'Percent Black', labels = scales::percent) + 
    ggtitle(city) + 
    ggthemes::theme_map() + 
    theme(legend.justification=c(0,0), legend.position=c(.7,0)) + 
    theme(plot.title = element_text(size=10, 
                                  vjust=1),
          legend.title = element_text(size = 8))

  ggsave(paste0('figs/', city, '_percent_black.pdf'), width = 10, height = 10, units = 'cm')

  tracts@data <- tracts@data %>% 
    mutate(non_black = Hispanic + Other + White + Asian)

  columns = c('non_black', 'Black')
  stuff <- compx::info_analysis(tracts, columns = columns, resolution = NULL, grid_tract = grid_tract, grid = grid)

  stuff$grid %>% 
    fortify(region = 'id') %>% 
    left_join(stuff$grid@data[,c('id', 'pop', 'info')], by = c('id' = 'id')) %>% 
    ggplot(aes(x = long, y = lat, fill =  info, group = group)) + 
    geom_polygon() + 
    viridis::scale_fill_viridis(name = expression(J[Y](x))) + 
    ggthemes::theme_map() +
    ggtitle(city) + 
    theme(legend.justification=c(0,0), legend.position=c(.7,0),
          legend.title = element_text(size = 8)) + 
    theme(plot.title = element_text(size=10, 
                                  vjust=1))

  # paste0('figs/', city, '_grid.pdf')
  ggsave(paste0('figs/', city, '_grid.pdf'), width = 10, height = 10, units = 'cm')

  percent_black = sum(tracts@data$Black) / sum(tracts@data$total)

  new_row <- data.frame(city = city, percent_black = percent_black, H_Y = stuff$H_Y, I_XY = stuff$I_XY, J_XY = stuff$J_XY)
  tab <- rbind(tab, new_row)
}

tab[,-1] <- round(tab[,-1],2)

tab %>% write_csv('figs/comparative_summary.csv')

### 2016-10-06 revision.tex

\usepackage{pc_writeup}
\usepackage{pc_math}
\usepackage[colorinlistoftodos]{todonotes}

@@

	\begin{figure}
		\includegraphics[width=\textwidth]{figs/checkerboard.png}
		\caption{Information and the checkerboard}. \label{fig:checkerboard}
	\end{figure}

@@

	\subsection{Illustrations}
		\subsubsection{Comparative Analysis: Detroit and Philadelphia}
			To illustrate these methods, we use them to compare the cities of Philadelphia and Detroit. Our data is blockgroup-level demographics from the American Community Survey \cite{CensusRace}. As we show, computing $H(Y)$, $I(X,Y)$, and $J(X,Y)$ together highlights similarities and differences between these cities. 
			\todo{We need to generate a nice summary plot here and then fill this section in.}
			\begin{figure}
				\caption{Spatial analysis of racial diversity in Detroit (left) and Philadelphia (right). Top: Percentage of black residents in Detroit (left) and Philadelphia (right). Middle: Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of $I(X,Y | X \in B_i)$. Bottom: Estimated contours of constant $I(X,Y | X \in B_i)$; high values correspond to spaces of demographic transition. Very bottom: Numerical summary, including the three information measures.} \label{fig:detroit_philly}
			\end{figure}

@@

			\begin{figure}
				\includegraphics[width=\textwidth]{figs/mutual_fisher.png}
				\caption{Comparative global diversity measures in major U.S. cities. Cities toward the bottom left are relatively spatially even. Cities toward the bottom right like Detroit and Atlanta are the most classically ``segregated'', with large, starkly separated monoracial regions. Cities toward the top right like Philadelphia and New York are complex ``patchworks'' of different, smaller regions.} \label{fig:mutual_fisher}
			\end{figure}

@@

	A useful place to start is with the property of ``additive spatial decomposition'' \cite{Reardon2004}. A diversity measure satisfies additive spatial decomposition when, if spatial units are aggregated into groups, the measure can be expressed as a sum of a ``within group'' term expressing variability within aggregations, and a ``between group'' term expressing variability across aggregations.

	\begin{itemize}
		\item Construct the algorithm. 
		\item Show algorithm results for one or two cities; show how knowing the clusters can lead to neat data analysis. 
		\item Relate algorithmic performance to suite of information measures above. Note that $J$ measures the ``natural fault lines'' where we would expect the algorithm to segment.  
	\end{itemize}

>>>>

\usepackage{pc_writeup}
\usepackage{pc_math}
\usepackage{csvsimple}
\usepackage[colorinlistoftodos]{todonotes}

@@

	\begin{figure}
		\includegraphics[width=\textwidth]{figs/checkerboard.pdf}
		\caption{Information and the checkerboard}. \label{fig:checkerboard}
	\end{figure}

@@

	\subsection{Illustrations}
		\subsubsection{Comparative Analysis: Detroit and Philadelphia}
			To illustrate these methods, we use them to compare the cities of Philadelphia and Detroit. Our data is blockgroup-level demographics from the American Community Survey \cite{CensusRace}. As we show, computing $H(Y)$, $I(X,Y)$, and $J(X,Y)$ together highlights similarities and differences between these cities. 
			\todo{We need to generate a nice summary plot here and then fill this section in.}
			\begin{figure}
				\includegraphics[width = .5\textwidth]{figs/Detroit_percent_black.pdf}
				\includegraphics[width = .5\textwidth]{figs/Detroit_grid.pdf} \\
				\includegraphics[width = .5\textwidth]{figs/Philadelphia_percent_black.pdf}
				\includegraphics[width = .5\textwidth]{figs/Philadelphia_grid.pdf} \\

				\centering
				\begin{tabular}{l | c c c c}
					\bfseries City & Percent Black & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$  \\\hline
					\csvreader[late after line=\\]{figs/comparative_summary.csv}{}
					{\csvcoli & \csvcolii & \csvcoliii & \csvcoliv & \csvcolv}
				\end{tabular}

				\caption{Spatial analysis of racial diversity in Detroit (top) and Philadelphia (middle). Left: Percentage of black residents in Detroit (left) and Philadelphia (right). Right: Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of $I(X,Y | X \in B_i)$. Bottom: Numerical summary, including the three information measures.} \label{fig:detroit_philly}
			\end{figure}

@@

			\begin{figure}
				\includegraphics[width=\textwidth]{figs/mutual_fisher.pdf}
				\caption{Comparative global diversity measures in major U.S. cities. Cities toward the bottom left are relatively spatially even. Cities toward the bottom right like Detroit and Atlanta are the most classically ``segregated'', with large, starkly separated monoracial regions. Cities toward the top right like Philadelphia and New York are complex ``patchworks'' of different, smaller regions.} \label{fig:mutual_fisher}
			\end{figure}

@@

	A useful place to start is with the property of ``additive spatial decomposition'' \cite{Reardon2004}. A diversity measure satisfies additive spatial decomposition when, if spatial units are aggregated into groups, the measure can be expressed as a sum of a ``within group'' term expressing variability within aggregations, and a ``between group'' term expressing variability across aggregations. The mutual information $I(X,Y)$ satisfies this property due to the following information-theoretic identity. Let $C$ be a random variable with the property that $H(C|X) = 0$; i.e. $X$ completely determines $C$. We can think of $C$ as group labels for the spatial locations $X$. Then, the chain rule of information implies that 
	\begin{equation}
		I(X,Y) = \underbrace{I(C,Y)}_{\text{between-group}} + \underbrace{I(X,Y|C)}_{\text{within-group}}. \label{eq:chain_rule}
	\end{equation}
	Equation \eqref{eq:chain_rule} expresses the following intuitive information decomposition: the information you have if I tell you first a resident's approximate residence $C$ and subsequently their exact residence $X$ is just the same as the information you would have if I simply told you $X$ directly.

	Equation \eqref{eq:chain_rule} also motivates a natural algorithm for identifying sociospatial structure. The core idea is that a clustering $C$ is ``good'' when it captures most of the intrinsic information in the system in the between-groups term $I(C,Y)$. This idea is analogous to between-group sum-of-squares maximization in classical ANOVA methods. We therefore seek to maximize $I(C,Y)$. In full generality, this is a challenging discrete optimization problem that may not be computationally tractable. We can, however, construct a ``greedy'' algorithm which, though suboptimal, leads to satisfactory results. At each stage of the algorithm, we consider the problem of choosing a pair of locales $\{i*, j*\}$ to aggregate together. A good choice of locales leads to minimal information loss, which we may quantify as follows. The mutual information may also be written: 
	\begin{align}
	 	I(X,Y) &= \sum_k p(X = k) D[p(Y|X = k\|p(Y)] 
	\end{align} 
	The contribution of two locales $i$ and $j$ to this sum is 
	\begin{align}
		I_{i,j}(X,Y) &\triangleq \sum_{k \in \{i,j\}} p(X = k)D[p(Y|X = k\|p(Y)] 
	\end{align}
	If we replace $i$ and $j$ with an aggregated local $i\cup j$, its contribution to the mutual information is 
	\begin{align}
		I_{i\cup j}(X,Y) = p(X \in \{i,j\}) D[p(Y|X \in \{i,j\}) \| p(Y)]\;.
	\end{align}
	The  loss associated with aggregating $i$ and $j$ is the difference between these two terms: 
	\begin{align}
		d(i,j) \triangleq I_{i,j}(X,Y) - I_{i \cup j}(X,Y).  
	\end{align}
	Like the KL divergence $D$, the function $d$ can be viewed as a measure of distance between probability distributions $p(Y|X = i)$ and $p(Y|X = j)$. Unlike the KL divergence, $d$ is symmetric in $i$ and $j$.

	\begin{itemize}
		\item Construct the algorithm. 
		\item Prove the triangle inequality
		\item Show algorithm results for one or two cities; show how knowing the clusters can lead to neat data analysis. 
		\item Relate algorithmic performance to suite of information measures above. Note that $J$ measures the ``natural fault lines'' where we would expect the algorithm to segment.  
	\end{itemize}
